Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-09 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: terse text on parspec formalism

Further expand the Kerr expression as:
%
\begin{subequations}
\begin{align}
\omega_{J} &= (1/M) \sum_{j = 0}^{N} \, \chi^{j} \omega^{(j)}_{k} \, \left( 1 + \gamma \delta \omega^{(j)}_{k} \right), \\
\tau_{J}   &= M     \sum_{j = 0}^{N} \, \chi^{j} \tau^{(j)}_{k}   \, \left( 1 + \gamma \delta \tau^{(j)}_{k} \right),
\end{align}
\label{eq:kerr_expansion}
\end{subequations}

Also,
%
\begin{equation}
\gamma
%
= \frac{\ell^{p}}{M_{\rm s}^{p}}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p}
\label{eq:def_gamma}
\end{equation}

@@

The prior probability distribution functions on the GR parameters are assumed to be uniform over the component masses, $(m_1, m_2)$, isotropically distributed on a sphere in the sky for the source location with $p(d_L) \propto d_L^2$, and isotropic on the binary orientation, $p(\iota, \psi, \phi_0) \propto \text{sin} \iota$. For the spins $(\chi_1, \chi_2)$, we assume a prior uniform and isotropic in the spin magnitudes and restrict ourselves to the z-components~\footnote{This spin-prior choice can be specified in \texttt{LALInference} using the option \texttt{alignedspin-zprior}.} The prior on $t_0$ is uniform and its range is informed by our detection pipelines.

@@

Finally, we use one of the approaches of Ref.~\cite{Perkins:2021mhb} to
compute the product in Eq.~\eqref{eq:cumulative_dist_ell}.
%
It method relies on first obtaining a kernel density estimator (KDE) for
the posterior distribution on $\ell$ for each individual event.
%
Next, we multiply these posteriors according to Eq.~\eqref{eq:cumulative_dist_ell}.
%
(See also Refs.~\cite{DelPozzo:2011pg,Cardenas-Avendano:2019zxd,Carullo:2021dui}.)
%
This approach based on the KDE can have difficulties in dealing with hard cut-offs in the
distribution, such as the one we impose at $\ell = 0$. We address this issue as
done in Ref.~\cite{Perkins:2021mhb}, by first doubling our number of samples as
$\{ s_i \} \cup \{ - s_i \}$, i.e., mirroring the samples across $\ell = 0$.
%
We then renormalize the associated KDE numerically in the range $\ell = [0, \infty]$.

@@

We presented an unified framework to combine the parametrized formalism introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model.
%
We showed through concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped upon the
theory-agnostic beyond-GR parameters of the {\sc ParSpec} formalism.
%
Together this allowed us to apply the \pSEOB~waveform model to test four
modified gravity theories (EdGB, dCS, cubic, and quartic
effective-field-theories of GR) using data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered the inspiral part of the GW signal alone.
%
There it is dCS which is unconstrained, while EdGB, in principle can be constrained, based on results for the closely
related shift-symmetric version of this theory.

>>>>

We expand the Kerr expression as:
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{J} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{J} \, \left( 1 + \gamma \delta \omega^{(n)}_{J} \right), \\
\tau_{J}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{J}   \, \left( 1 + \gamma \delta \tau^{(n)}_{J} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
remnant black hole; $\omega_{K}^{n}$, $\tau_{K}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{K}^{n}$, $\delta\tau_{K}^{n}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p}
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of lenght, which sets the
lenghtscale in which beyond-GR modification become important and the exponent $p$
which is related to how the beyond-GR modifications are included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also in terms of the detector-frame mass $M$ through
a the redshift $z$.

In principle, a modification to GR would also affect $M$ and $\chi$ and the expansion should be written
in terms of non-GR mass and spin, say $\bar{M}$ and $\bar{\chi}$. If we assume that the beyond-GR corrections
are included perturbatively, the modifications to the BH mass and spin can be absorved into the deviations
parameters $\delta\omega$ and $\delta\tau$.
%
This means we can identify the $M$ and $\chi$ with their corresponding GR values.

\hs{TODO: in figures and tables use $N_{\rm max} = 0$ ($N_{\rm max} = 1$) instead of $j=0$ ($j=0,1$).}
\hs{TODO: say that we check a posteriori if our PE runs agree with this.}

@@

The prior probability distribution functions on the GR parameters are assumed to be uniform over the component masses, $(m_1, m_2)$, isotropically distributed on a sphere in the sky for the source location with $p(d_L) \propto d_L^2$, and isotropic on the binary orientation, $p(\iota, \psi, \phi_0) \propto \text{sin} \iota$. For the spins $(\chi_1, \chi_2)$, we assume a prior uniform and isotropic in the spin magnitudes and restrict ourselves to the $z$-components.
This spin-prior choice can be specified in \texttt{LALInference} using the option \texttt{alignedspin-zprior}.
%
The prior on $t_0$ is uniform and its range is informed by our detection pipelines.

@@

Finally, we use one of the approaches of Ref.~\cite{Perkins:2021mhb} to
compute the product in Eq.~\eqref{eq:cumulative_dist_ell}.
%
It method relies on first obtaining a kernel density estimator (KDE) for
the posterior distribution on $\ell$ for each individual event.
%
(See also Refs.~\cite{DelPozzo:2011pg,Cardenas-Avendano:2019zxd,Carullo:2021dui}.)
%
Next, we multiply these posteriors according to Eq.~\eqref{eq:cumulative_dist_ell}.
%
This approach based on the KDE can have difficulties in dealing with hard cut-offs in the
distribution, such as the one we impose at $\ell = 0$. We address this issue as
done in Ref.~\cite{Perkins:2021mhb}, by first doubling our number of samples as
$\{ s_i \} \cup \{ - s_i \}$, i.e., mirroring the samples across $\ell = 0$.
%
We then renormalize the associated KDE numerically in the range $\ell = [0, \infty]$.

@@

We presented an unified framework to combine the parametrized formalism introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model.
%
We showed through concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped upon the
theory-agnostic beyond-GR parameters of the {\sc ParSpec} formalism.
%
Together this allowed us to apply the \pSEOB~waveform model to test four
modified gravity theories (EdGB, dCS, cubic, and quartic
effective-field-theories of GR) using data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered
the inspiral part of the GW signal alone.
%
In that case, it is dCS which is unconstrained, while EdGB, in principle can be
constrained, based on results for the closely related shift-symmetric version
of this theory.
